High-order cognition is supported by information-rich but compressible brain activity patterns

### 2023-03-17 paper/main.tex
final round of edits before submission

We take as a given that brain activity is highly flexible: our brains can
exhibit nearly infinite varieties of activity patterns. This flexibility implies that our
brains activity patterns are highly information rich. However, brain activity
patterns are also highly structured. For example, full-brain correlation
matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
stability suggests that our brains' activity patterns are at least partially
constrained, for example by anatomical, external, or internal factors.
Constraints on brain activity that limit its flexibility decrease
expressiveness (i.e., its information rate). However, constraints on brain
activity also increase its robustness to noise (e.g., ``missing'' or corrupted
signals may be partially recovered). For example, recent work has shown that
full-brain activity patterns may be reliably recovered from only a relatively
small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
robustness property suggests that the relevant signal (e.g., underlying factors
that have some influence over brain activity patterns) are compressible.

@@

  \caption{\textbf{Information content and compressibility.} \textbf{A.
  Variance explained for two images.} We applied principal components analysis
  to a photograph and drawing, treating the rows of the images as
  ``observations.'' Across columns, we identified the numbers of components
  required to explain 100\%, 95\%, 75\%, or 50\% of the cumulative variance in
  each image (the 100\% columns denote the original images). The numbers of
  components are indicate in parentheses, and the resulting ``compressed''
  images are displayed. \textbf{B. Representing two images with different
  numbers of components.} Using the same principal component decompositions as
  in Panel A, we computed the cumulative proportion of variance explained with
  380 (original images), 20, 4, or 2 components. \textbf{C. Cumulative variance
  explained versus number of components.} For the images displayed in Panels A
  and B, we plot the cumulative proportion of variance explained as a function
  of the number of components used to represent each image. \textbf{D.
  Information rate and compressibility.} Across multiple images, the
  information rate (i.e., the amount of information contained in each image;
  horizontal axis) is high when each individual pixel provides information that
  cannot be inferred from other pixels. High-information rate images tend to be
  high-resolution, and low-information rate images tend to be low-resolution.
  Compressibility is related to the difference between the information required
  to specify the original versus compressed images (vertical axis). Highly
  compressible images often contain predictable structure (redundancies) that
  can be leveraged to represent the images much more efficiently than in their
  original feature spaces.}

@@

We analyzed a dataset collected by \cite{SimoEtal16} that comprised four
experimental conditions. These conditions exposed participants to stimuli that
systematically varied in cognitive engagement. In the \textit{intact}
experimental condition, participants listened to an audio recording of a
10-minute story. In the \textit{paragraph}-scrambled experimental condition,
partipants listened to a temporally scrambled version of the story, where the
paragraphs occurred out of order, but where the same set of paragraphs was
presented over the entire listening interval. All participants in this
condition experienced the scrambled paragraphs in the same order. In the
\textit{word}-scrambled experimental condition, participants listened to a
temporally scrambled version of the story, where the words occurred in a random
order. Again, all participants in this condition experienced the scrambled
words in the same order. Finally, in the \textit{rest} experimental condition,
participants lay in the scanner with no overt stimulus, while keeping their
eyes open and blinking as needed. This public dataset provided a convenient
means for testing our hypothesis that different levels of cognitive processing
and engagement affect how informative and compressible the associated brain
patterns are.

@@

To evaluate the relation between informativeness and compressibiilty for brain
activity from each experimental condition, we trained a series of
across-participant temporal decoders on compressed representations of the data.
Figure~\ref{fig:inflection}A displays the decoding accuracy as a function of
the number of principal components used to represent the data (also see
Fig.~\varExplained). Several patterns were revealed by the analysis. First, in
general (i.e., across experimental conditions), decoding accuracy tends to
improve as the number of components are increased. However, decoding accuracy
peaked at higher levels for experimental conditions that exposed participants
to cognitively richer stimuli (Fig.~\ref{fig:inflection}D). The peak decoding
accuracy was highest for the ``intact'' condition (versus paragraph: $t(99) =
35.205, p < 0.001)$; versus word: $t(99) = 43.172, p < 0.001)$; versus rest:
$t(99) = 81.361, p < 0.001)$), next highest for the ``paragraph'' condition
(versus word: $t(99) = 6.243, p < 0.001)$; versus rest: $t(99) = 50.748, p <
0.001)$), and next highest for the ``word'' condition (versus rest: $t(99) =
48.791, p < 0.001)$). This ordering implies that cognitively richer conditions
evoke more stable brain activity patterns across people.

The cognitively richer conditions also displayed steeper initial slopes. For
example, the intact condition decoders reached an arbitrarily chosen threshold
of 5\% accuracy using fewer components than the paragraph condition decoders
($t(99) = -7.429, p < 0.001$) or word condition decoders ($t(99) = -7.300, p <
0.001$), and decoding accuracy never exceeded 5\% for the rest condition. This
suggests that brain activity patterns evoked by cognitively richer conditions
are more compressible, such that representing the data using the same number of
principal components provides more information to the temporal decoders
(Figs.~\ref{fig:inflection}B, C). Taken together, as shown in
Figure~\ref{fig:inflection}E, we found that brain activity patterns evoked by
cognitively richer conditions tended to be both more informative (i.e.,
associated with higher decoding accuracies) \textit{and} more compressible
(i.e., requiring fewer components to achieve the 5\% accuracy threshold).

@@

\caption{\textbf{Changes in decoding accuracy and compression over time.}
\textbf{A. Decoding accuracy by number of components, by story segment.} Each
family of curves is plotted in the same format as Figure~\ref{fig:inflection}A
but reflects data only from one quarter of the dataset. \textbf{B.
Informativeness versus compressibility by condition and segment.} Each scatter
plot is in the same format as Figure~\ref{fig:inflection}E, but reflects data
only from one quarter of the dataset. \textbf{C. Change in decoding accuracy
over time, by number of components.} For each number of components ($x$-axis)
and condition (color), we fit a regression line to the decoding accuracies
obtained for the first, second, third, and fourth quarters of the dataset
(corresponding to the columns of Panels A and B). The $y$-axis denotes the
slopes of the regression lines. All error ribbons denote bootstrap-estimated
95\% confidence intervals.}

@@

  \caption{\textbf{Network-specific decoding accuracy and compression.}
  \textbf{A. Network parcellation map.} The glass brain displays the location
  of each of seven networks identified by \cite{YeoEtal11}. \textbf{B. Node
  labels.} We assigned network labels to each of 700 Hierarchical Topographic
  Factor Analysis-derived nodes (see \textit{Hierarchical topographic factor
  analysis (HTFA)}). For each node, we evaluated its radial basis function
  (RBF) at the locations of every voxel in the parcellation map displayed in
  Panel A. We summed up these RBF weights separately for the voxels in each
  network. We defined each node's network label as its highest-weighted network
  across all voxels. \textbf{C. Per-network node counts.} The bar plot displays
  the total number of HTFA-derived nodes assigned to each network. \textbf{D.
  Informativeness versus compressibility by network and condition.} Each dot
  corresponds to a single cross validation run. The dots' coordinates denote
  the minimum proportion of components (relative to the number of nodes in the
  given network) required to achieve at least 5\% corrected decoding accuracy
  ($x$-coordinate; for the rest condition these proportions are estimated using
  the peak decoding accuracies) and peak (corrected) decoding accuracy
  ($y$-coordinate). We repeated the analysis separately for each network
  (color) and experimental condition (column). \textbf{E. Informativeness by
  network.} For each network, sorted (roughly) from lower-order networks on the
  left to higher-order networks on the right, the curves display the peak
  (corrected) decoding accuracy for each experimental condition (color).
  \textbf{F. Compressibility by network.} For each network, the curves display
  the proportion of components (relative to the total possible number of
  components displayed in Panel C) required to achieve at least 5\% decoding
  accuracy (or, for the rest condition, peak decoding accuracy, as described
  above). Error ribbons in Panels E and F denote 95\% bootstrap-estimated
  confidence intervals.}

@@

We also wondered how informativeness and compressibility in the different
experimental conditions might vary across brain networks. We used a network
parcellation identified by \cite{YeoEtal11} to segment the brain into seven
distinct networks. The networks can be sorted (roughly) in order from
lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
visual, somatomotor, dorsal attention, ventral attention, limbic,
frontoparietal, and default mode. Next, we computed decoding curves separately
for the activity patterns within each network and identified each network's
inflection point, for each experimental condition. Moving from low-order
networks to higher-order networks, we found that decoding accuracy tended to
increase, particularly in the higer-order experimental conditions
(Fig.~\ref{fig:networks}D, E). This suggests that higher-order networks may
carry more content-relevant or stimulus-driven ``information.'' We found no
clear trends in the proportions of components required to achieve 5\% decoding
accuracy across networks or conditions (Fig.~\ref{fig:networks}F).

@@

\caption{\textbf{Neurosynth topic weightings by component.} We used a reverse
inference procedure (see \textit{Reverse inference}) to estimate the
correspondences between brain images and a set of 80 topics derived from the
Neurosynth database of neuroimaging articles~\citep{RubiEtal17}. \textbf{A.
Topic correlations by component.} For each of the top five highest-weighted
principal components (columns) derived from each experimental condition
(colors), the radar plots display the correlations between the component images
and the per-topic images derived for each topic (topic identities are denoted
by the blue dots; a legend for the topic labels is in the lower right of the
figure). An annotated list of the top-weighted topics for each component and
condition may be found in Figure~\topTerms~and the top-weighted terms for each
topic may be found in Table~\topics. \textbf{B. Topic correlations averaged
across components.} The radar plot is in the same format as the plots in Panel
A, but here we display the per-condition average correlations across the top
five components (for each condition) reflected in Panel A. \textbf{C. Component
images.} Each plot diplays a right sagittal view of a glass brain projection of
the top five principal components (columns) for each experimental conditions
(rows). Additional projections for each component may be found in
Figure~\componentBrains.}

@@

To gain further insights into which brain functions might be most closely
associated with the top-weighted components from each experimental condition,
we manually grouped each Neurosynth-derived topic into a smaller set of core
cognitive functions. Separately for each component, we computed the average
weightings across all topics that were tagged as being associated with each of
these cognitive functions (Figs.~\ref{fig:neurosynth-summary}A,~\neurosynthFull
A). To help visualize these associations, we used the patterns of associations
for each component to construct graphs whose nodes were experimental conditions
and cognitive functions (Figs.~\ref{fig:neurosynth-summary}B,~\neurosynthFull
B). We also computed correlations between the sets of per-topic weightings from
each of the top-weighted components from each experimental condition
(Fig.~\ref{fig:neurosynth-summary}C, D) and between the brain maps for each
condition's components (Fig.~\neurosynthFull C, D). Taken together, we found
that each component appeared to weight on a fundamental set of cognitive
functions that varied by experimental condition. For example, the top principal
components from every condition weighted similarly on the full set of
Neurosynth topics (Fig.~\ref{fig:neurosynth-pca}A) and cognitive functions
(Figs.~\ref{fig:neurosynth-summary}A, B and~\neurosynthFull~A, B), suggesting
that these components might reflect a set of functions or activity patterns
that are common across all conditions. The second components' weightings were
similar across the intact, paragraph, and rest conditions (highest-weighted
functions: cognitive control, memory, social cognition, and resting state), but
different for the word condition (highest-weighted functions: sensory
perception and cognitive control). The fourth components' weighting 
grouped the paragraph and word conditions (highest-weighted functions: memory,
language processing, and cognitive control) and the intact and rest conditions
(highest-weighted functions: emotion, social cognition). We also used
ChatGPT~\citep{ChatGPT} to sort the list of manually tagged cognitive functions
from lowest-level to highest-level (Tab.~\topicTags,
Fig.~\ref{fig:neurosynth-summary}E; also see \textit{Ranking cognititive
processes}). We found that higher-level functions tended to be weighted on more
heavily by top components from the intact and paragraph conditions than
lower-level functions. The top components from the word condition showed the
opposite tendancy, whereby \textit{lower}-level functions tended to be weighted
on more heavily than higher-level functions. The components from the rest
condition showed almost no differences in the weights associated with
high-level versus low-level functions.

@@

\caption{\textbf{Summary of functions associated with top-weighted components
by condition.} \textbf{A. Top-weighted topics by condition.} Here we display
per-condition (rows, indicated by colored dots) topic correlations, averaged
across topics that pertain to each of several broad cognitive functions
(columns within each sub-panel, indicated by colored dots). Each sub-panel
reflects correlations for the components indicated in the panel titles. A
legend for the condition and several core cognitive function classifications is
displayed in the lower right of the figure. Table~\topics~provides a list of
each topic's top-weighted terms, along with each topic's manually labeled
cognitive classification. A full list of the topics most highly associated with
each component may be found in Figure~\topTerms. \textbf{B. Associations
between per-condition components and cognitive functions.} The network plots
denote positive average correlations between the component images for each
condition (gray-outlined dots on the left sides of each network; colors denote
conditions) and topic-specific brain maps associated with each indicated
cognitive function (black-outlined dots on the right sides of each network;
colors denote cognitive functions). The line thicknesses are proportional to
the correlation values (correlation coefficients are noted in the heatmaps in
Panel A). \textbf{C. Correlations between each principal component, by
condition.} The heatmaps display the correlations between the brain maps
(Fig.~\componentBrains) for each principal component (sub-panel), across each
pair of conditions (rows and columns of each sub-panel's matrix, indicated by
colored dots). \textbf{D. Associations between per-condition components, by
component.} Each sub-panel's network plot summarizes the pattern of correlations
between the $n$\textsuperscript{th} top-weighted principal components
(sub-panel) for each experimental condition (gray-outlined dots). The line
thicknesses are proportional to the correlation values (correlation
coefficients are noted in the heatmaps in Panel C). \textbf{E. Change in
weights by condition.} The bar heights reflect the slopes of regression lines,
fit separately to the top five components from each condition, between the
ChatGPT-derived ``rank'' of each cognitive classification and the correlations
between the component and topic maps associated with cognitive processes at the
given rank (see \textit{Ranking cognitive processes}).  Also see
Fig.~\neurosynthFull~for additional information.}

@@

Our explorations of informativeness and compressibility are related to a much
broader literature on the correlational and causal structure of brain activity
patterns~\citep{PretEtal17, OwenEtal21, RogeEtal07, RubiSpor10, SizeEtal18,
SmitEtal13b, SmitEtal13c, SrinEtal07, TomaVolk11, YeoEtal11, AdacEtal12,
BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04, DhamEtal08,
KorzEtal08, BrovEtal04}. Correlations or causal associations between different
brain regions simultaneously imply that full-brain activity patterns will be
compressible and also that those activity patterns will contain redundancies.
For example, the extent to which activity patterns at one brain area can be
inferred or predicted from activity patterns at other
areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21}, reflects overlap in the
information available in or represented by those brain areas. If brain patterns
in one area are recoverable using brain patterns in another area, then a
``signal'' used to convey the activity patterns could be compressed by removing
the recoverable activity. Predictable (and therefore redundant) brain activity
patterns are also more robust to signal corruption. For example, even if the
activity patterns at one region are unreadable or unreliable at a given moment,
that unreliability could be compensated for by other regions' activity patterns
that were predictive of the unreliable region.

@@

Cognitive neuroscientists are still grappling with basic questions about the
fundamental ``rules'' describing how our brains respond, and about how brain
activity patterns and the associated underlying cognitive representations and
computations are linked. We identified two aspects of brain activity patterns,
informativeness and compressibiilty, that appear to change systematically with
task demands and across brain networks. Our work helps to clarify how the
``neural code'' might be structured, and how the code might vary across tasks
and brain areas.

@@

Next, we considered 80 topics estimated using Latent Dirichlet
Allocation~\citep{BleiEtal03} applied to 9,204 functional neuroimaging articles
in the Neurosynth database~\citep{RubiEtal17}. The topics, as well as
associated brain maps identified using Neurosynth, were identified and reported
in several prior studies~\citep{FoxEtal14, SulEtal17, ChenEtal20}. The topic
labels for each topic were generated automatically with the following ChatGPT
(chat.openai.com) prompt: ``Please help me come up with intuitive labels for
topic topics I found by fitting a topic model to thousands of neuroscience and
psychology articles. I'll paste in the top 10 highest-weighted words for each
topic, and I'd like you to respond with a suggested label. For each topic,
please respond with just the topic label and no other formatting or text. Here
are the next topic's top words:'' followed by a comma-separated list of the
given topic's top-weighted words reflected in the Table~\topics. For some
topics, ChatGPT responded with a longer-form response rather than a concise
topic label. In these instances, on a case-by-case basis, we used a second
follow-up prompt to achieve the given topic's label: ``Could you please come up
with a more concise label for that topic?''. We then manually identified a set
of 11 cognitive labels that were intended to encapsulate a representative range
of widely studied low-level and high-level cognitive functions. In choosing the
set of cognitive labels, we jointly considered each topic's ChatGPT-derived
topic label, along with the top-weighted words for the topic. We attempted to
generate a concise set of labels that still spanned the full set of cognitive
functions reflected across the 80 topics. Topics that appeared unrelated to
specific cognitive functions (e.g., topics related to specific methods or
clinical themes) are designated with dashes in Table~\topics.

Finally, following an approach used in several prior studies~\citep{FoxEtal14,
SulEtal17, ChenEtal20} we treated the correlation between a given component's
brain map and each topic's brain map as a reflection of how much the component
was reflective of the given topic. This resulted in a set of 80 ``weights''
(correlation coefficients) for each component's brain map, with one weight per
Neurosynth-derived topic.

@@

Conceptulization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
Implementation: L.L.W.O. Analysis: L.L.W.O. Writing, Reviewing, and Editing:
J.R.M. and L.L.W.O. Supervision: J.R.M.

>>>>

We take as a given that brain activity is highly flexible: our brains can
exhibit nearly infinite varieties of activity patterns. This flexibility
implies that our brains' activity patterns are highly information rich. However,
brain activity patterns are also highly structured. For example, full-brain
correlation matrices are stable within~\citep{FinnEtal15, FinnEtal17,
GratEtal18} and across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14}
individuals. This stability suggests that our brains' activity patterns are at
least partially constrained, for example by anatomical, external, or internal
factors. Constraints on brain activity that limit its flexibility decrease
expressiveness (i.e., its information rate). However, constraints on brain
activity also increase its robustness to noise (e.g., ``missing'' or corrupted
signals may be partially recovered). For example, recent work has shown that
full-brain activity patterns may be reliably recovered from only a relatively
small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
robustness property suggests that the relevant signal (e.g., underlying factors
that have some influence over brain activity patterns) are compressible.

@@

  \caption{\textbf{Information content and compressibility.} \textbf{A.
  Variance explained for two images.} We applied principal components analysis
  to a photograph and drawing, treating the rows of the images as
  ``observations.'' Across columns, we identified the numbers of components
  required to explain 100\%, 95\%, 75\%, or 50\% of the cumulative variance in
  each image (the 100\% columns denote the original images). The numbers of
  components are indicated in parentheses, and the resulting ``compressed''
  images are displayed. \textbf{B. Representing two images with different
  numbers of components.} Using the same principal component decompositions as
  in Panel A, we computed the cumulative proportion of variance explained with
  380 (original images), 20, 4, or 2 components. \textbf{C. Cumulative variance
  explained versus number of components.} For the images displayed in Panels A
  and B, we plot the cumulative proportion of variance explained as a function
  of the number of components used to represent each image. \textbf{D.
  Information rate and compressibility.} Across multiple images, the
  information rate (i.e., the amount of information contained in each image;
  horizontal axis) is high when each individual pixel provides information that
  cannot be inferred from other pixels. High-information rate images tend to be
  high-resolution, and low-information rate images tend to be low-resolution.
  Compressibility is related to the difference between the information required
  to specify the original versus compressed images (vertical axis). Highly
  compressible images often contain predictable structure (redundancies) that
  can be leveraged to represent the images much more efficiently than in their
  original feature spaces.}

@@

We analyzed a dataset collected by \cite{SimoEtal16} that comprised four
experimental conditions. These conditions exposed participants to stimuli that
systematically varied in cognitive engagement. In the \textit{intact}
experimental condition, participants listened to an audio recording of a
10-minute Moth Radio Hour story, \textit{Pie Man}, by Jim O'Grady. In the
\textit{paragraph}-scrambled experimental condition, participants listened to a
temporally scrambled version of the story, where the paragraphs occurred out of
order, but where the same set of paragraphs was presented over the entire
listening interval. All participants in this condition experienced the
scrambled paragraphs in the same order. In the \textit{word}-scrambled
experimental condition, participants listened to a temporally scrambled version
of the story, where the words occurred in a random order. Again, all
participants in this condition experienced the scrambled words in the same
order. Finally, in the \textit{rest} experimental condition, participants lay
in the scanner with no overt stimulus, while keeping their eyes open and
blinking as needed. This public dataset provided a convenient means for testing
our hypothesis that different levels of cognitive processing and engagement
affect how informative and compressible the associated brain patterns are.

@@

To evaluate the relation between informativeness and compressibility for brain
activity from each experimental condition, we trained a series of
across-participant temporal decoders on compressed representations of the data.
Figure~\ref{fig:inflection}A displays the decoding accuracy as a function of
the number of principal components used to represent the data (also see
Fig.~\varExplained). Several patterns were revealed by the analysis. First, in
general (i.e., across experimental conditions), decoding accuracy tends to
improve as the number of components are increased. However, decoding accuracy
peaked at higher levels for experimental conditions that exposed participants
to cognitively richer stimuli (Fig.~\ref{fig:inflection}D). The peak decoding
accuracy was highest for the ``intact'' condition (versus paragraph: $t(99) =
35.205, p < 0.001)$; versus word: $t(99) = 43.172, p < 0.001)$; versus rest:
$t(99) = 81.361, p < 0.001)$), next highest for the ``paragraph'' condition
(versus word: $t(99) = 6.243, p < 0.001)$; versus rest: $t(99) = 50.748, p <
0.001)$), and next highest for the ``word'' condition (versus rest: $t(99) =
48.791, p < 0.001)$). This ordering implies that cognitively richer conditions
evoke more stable brain activity patterns across people.

The cognitively richer conditions also displayed steeper initial slopes. For
example, the intact condition decoders reached an arbitrarily chosen threshold
of 5\% accuracy using fewer components than the paragraph condition decoders
($t(99) = -7.429, p < 0.001$) or word condition decoders ($t(99) = -7.300, p <
0.001$), and decoding accuracy never exceeded 5\% for the rest condition. This
suggests that brain activity patterns evoked by cognitively richer conditions
are more compressible, such that representing the data using the same number of
principal components provides more information to the temporal decoders
(Figs.~\ref{fig:inflection}B, C). Taken together, as shown in
Figure~\ref{fig:inflection}E, we found that brain activity patterns evoked by
cognitively richer conditions tended to be both more informative (i.e.,
associated with higher peak decoding accuracies) \textit{and} more compressible
(i.e., requiring fewer components to achieve the 5\% accuracy threshold).

@@

\caption{\textbf{Changes in decoding accuracy and compression over time.}
\textbf{A. Decoding accuracy by number of components, by story segment.} Each
family of curves is plotted in the same format as Figure~\ref{fig:inflection}A
but reflects data only from one quarter of the dataset. \textbf{B.
Informativeness versus compressibility by condition and segment.} Each scatter
plot is in the same format as Figure~\ref{fig:inflection}E, but reflects data
only from one quarter of the dataset. \textbf{C. Change in decoding accuracy
over time, by number of components.} For each number of components ($x$-axis)
and condition (color), we fit a regression line to the decoding accuracies
obtained for the first, second, third, and fourth quarters of the dataset
(corresponding to the columns of Panels A and B). The $y$-axis denotes the
slopes of the regression lines. The black vertical line marks $k = 20$ components,
as referenced in the main text.  All error ribbons denote bootstrap-estimated
95\% confidence intervals.}

@@

  \caption{\textbf{Network-specific decoding accuracy and compression.}
  \textbf{A. Network parcellation map.} The glass brain displays the location
  of each of seven networks identified by \cite{YeoEtal11}. \textbf{B. Node
  labels.} We assigned network labels to each of 700 Hierarchical Topographic
  Factor Analysis-derived nodes (see \textit{Hierarchical topographic factor
  analysis (HTFA)}). For each node, we evaluated its radial basis function
  (RBF) at the locations of every voxel in the parcellation map displayed in
  Panel A. We summed up these RBF weights separately for the voxels in each
  network. We defined each node's network label as its highest-weighted network
  across all voxels. \textbf{C. Per-network node counts.} The bar plot displays
  the total number of HTFA-derived nodes assigned to each network. \textbf{D.
  Informativeness versus compressibility by network and condition.} Each dot
  corresponds to a single cross validation run. The dots' coordinates denote
  the minimum proportion of components (relative to the number of nodes in the
  given network) required to achieve at least 5\% corrected decoding accuracy
  ($x$-coordinate; for the rest condition these proportions are estimated using
  the peak decoding accuracies) and peak (corrected) decoding accuracy
  ($y$-coordinate). We repeated the analysis separately for each network
  (color) and experimental condition (column). \textbf{E. Informativeness by
  network.} For each network, sorted (roughly) from lower-order networks on the
  left to higher-order networks on the right, the curves display the peak
  (corrected) decoding accuracy for each experimental condition (color).
  \textbf{F. Compressibility by network.} For each network, the curves display
  the proportion of components (relative to the total possible number of
  components displayed in Panel C) required to achieve at least 5\% decoding
  accuracy (or, for the rest condition, peak decoding accuracy, as described
  above). Error ribbons in Panels E and F denote bootstrap-estimated
  95\% confidence intervals.}

@@

We also wondered how informativeness and compressibility in the different
experimental conditions might vary across brain networks. We used a network
parcellation identified by \cite{YeoEtal11} to segment the brain into seven
distinct networks. The networks can be sorted (roughly) in order from
lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
visual, somatomotor, dorsal attention, ventral attention, limbic,
frontoparietal, and default mode. Next, we computed decoding curves separately
for the activity patterns within each network and identified each network's
inflection point, for each experimental condition. Moving from low-order
networks to higher-order networks, we found that decoding accuracy tended to
increase, particularly in the higher-level experimental conditions
(Fig.~\ref{fig:networks}D, E). This suggests that higher-order networks may
carry more content-relevant or stimulus-driven ``information.'' We found no
clear trends in the proportions of components required to achieve 5\% decoding
accuracy across networks or conditions (Fig.~\ref{fig:networks}F).

@@

\caption{\textbf{Neurosynth topic weightings by component.} We used a reverse
inference procedure (see \textit{Reverse inference}) to estimate the
correspondences between brain images and a set of 80 topics derived from the
Neurosynth database of neuroimaging articles~\citep{RubiEtal17}. \textbf{A.
Topic correlations by component.} For each of the top five highest-weighted
principal components (columns) derived from each experimental condition
(colors), the radar plots display the correlations between the component images
and the per-topic images derived for each topic (topic identities are denoted
by the blue dots; a legend for the topic labels is in the lower right of the
figure). An annotated list of the top-weighted topics for each component and
condition may be found in Figure~\topTerms~and the top-weighted terms for each
topic may be found in Table~\topics. \textbf{B. Topic correlations averaged
across components.} The radar plot is in the same format as the plots in Panel
A, but here we display the per-condition average correlations across the top
five components (for each condition) reflected in Panel A. \textbf{C. Component
images.} Each plot displays a right sagittal view of a glass brain projection of
the top five principal components (columns) for each experimental condition
(rows). Additional projections for each component may be found in
Figure~\componentBrains.}

@@

To gain further insights into which brain functions might be most closely
associated with the top-weighted components from each experimental condition,
we manually grouped each Neurosynth-derived topic into a smaller set of core
cognitive functions. Separately for each component, we computed the average
weightings across all topics that were tagged as being associated with each of
these cognitive functions (Figs.~\ref{fig:neurosynth-summary}A,~\neurosynthFull
A). To help visualize these associations, we used the patterns of associations
for each component to construct graphs whose nodes were experimental conditions
and cognitive functions (Figs.~\ref{fig:neurosynth-summary}B,~\neurosynthFull
B). We also computed correlations between the sets of per-topic weightings from
each of the top-weighted components from each experimental condition
(Fig.~\ref{fig:neurosynth-summary}C, D) and between the brain maps for each
condition's components (Fig.~\neurosynthFull C, D). Taken together, we found
that each component appeared to weight on a fundamental set of cognitive
functions that varied by experimental condition. For example, the top principal
components from every condition weighted similarly (across conditions) on the
full set of Neurosynth topics (Fig.~\ref{fig:neurosynth-pca}A) and cognitive
functions (Figs.~\ref{fig:neurosynth-summary}A, B and~\neurosynthFull~A, B),
suggesting that these components might reflect a set of functions or activity
patterns that are common across all conditions. The second components'
weightings were similar across the intact, paragraph, and rest conditions
(highest-weighted functions: cognitive control, memory, social cognition, and
resting state), but different for the word condition (highest-weighted
functions: sensory perception and cognitive control). The fourth components'
weighting grouped the paragraph and word conditions (highest-weighted
functions: memory, language processing, and cognitive control) and the intact
and rest conditions (highest-weighted functions: emotion, social cognition). We
also used ChatGPT~\citep{ChatGPT} to sort the list of manually tagged cognitive
functions from lowest-level to highest-level (Tab.~\topicTags,
Fig.~\ref{fig:neurosynth-summary}E; also see \textit{Ranking cognitive
processes}). We found that higher-level functions tended to be weighted on more
heavily by top components from the intact and paragraph conditions than
lower-level functions. The top components from the word condition showed the
opposite tendency, whereby \textit{lower}-level functions tended to be weighted
on more heavily than higher-level functions. The components from the rest
condition showed almost no differences in the weights associated with
high-level versus low-level functions.

@@

\caption{\textbf{Summary of functions associated with top-weighted components
by condition.} \textbf{A. Top-weighted topics by condition.} Here we display
per-condition (rows, indicated by colored dots) topic correlations, averaged
across topics that pertain to each of several broad cognitive functions
(columns within each sub-panel, indicated by colored dots). Each sub-panel
reflects correlations for the components indicated in the panel titles. A
legend for the condition and several core cognitive function classifications is
displayed in the lower right of the figure. Table~\topics~provides a list of
each topic's top-weighted terms, along with each topic's manually labeled
cognitive classification. A full list of the topics most highly associated with
each component may be found in Figure~\topTerms. \textbf{B. Associations
between per-condition components and cognitive functions.} The network plots
denote positive average correlations between the component images for each
condition (gray-outlined dots on the left sides of each network; colors denote
conditions) and topic-specific brain maps associated with each indicated
cognitive function (black-outlined dots on the right sides of each network;
colors denote cognitive functions). The line thicknesses are proportional to
the correlation values (correlation coefficients are noted in the heat maps in
Panel A). \textbf{C. Correlations between each principal component, by
condition.} The heat maps display the correlations between the brain maps
(Fig.~\componentBrains) for each principal component (sub-panel), across each
pair of conditions (rows and columns of each sub-panel's matrix, indicated by
colored dots). \textbf{D. Associations between per-condition components, by
component.} Each sub-panel's network plot summarizes the pattern of correlations
between the $n$\textsuperscript{th} top-weighted principal components
(sub-panel) for each experimental condition (gray-outlined dots). The line
thicknesses are proportional to the correlation values (correlation
coefficients are noted in the heat maps in Panel C). \textbf{E. Change in
weights by condition.} The bar heights reflect the slopes of regression lines,
fit separately to the top five components from each condition, between the
ChatGPT-derived ``rank'' of each cognitive classification and the correlations
between the component and topic maps associated with cognitive processes at the
given rank (see \textit{Ranking cognitive processes}).  Also see
Fig.~\neurosynthFull~for additional information.}

@@

Our explorations of informativeness and compressibility are related to a much
broader literature on the correlational and causal structure of brain activity
patterns and networks~\citep{PretEtal17, OwenEtal21, RogeEtal07, RubiSpor10,
SizeEtal18, SmitEtal13b, SmitEtal13c, SrinEtal07, TomaVolk11, YeoEtal11,
AdacEtal12, BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04,
DhamEtal08, KorzEtal08, BrovEtal04, LynnBass21}. Correlations or causal
associations between different brain regions simultaneously imply that
full-brain activity patterns will be compressible and also that those activity
patterns will contain redundancies. For example, the extent to which activity
patterns at one brain area can be inferred or predicted from activity patterns
at other areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21}, reflects overlap in the
information available in or represented by those brain areas. If brain patterns
in one area are recoverable using brain patterns in another area, then a
``signal'' used to convey the activity patterns could be compressed by removing
the recoverable activity. Predictable (and therefore redundant) brain activity
patterns are also more robust to signal corruption. For example, even if the
activity patterns at one region are unreadable or unreliable at a given moment,
that unreliability could be compensated for by other regions' activity patterns
that were predictive of the unreliable region.

@@

Cognitive neuroscientists are still grappling with basic questions about the
fundamental ``rules'' describing how our brains respond, and about how brain
activity patterns and the associated underlying cognitive representations and
computations are linked. We identified two aspects of brain activity patterns,
informativeness and compressibility, that appear to change systematically with
task demands and across brain networks. Our work helps to clarify how the
``neural code'' might be structured, and how the code might vary across tasks
and brain areas.

@@

Next, we considered 80 topics estimated using Latent Dirichlet
Allocation~\citep{BleiEtal03} applied to 9,204 functional neuroimaging articles
in the Neurosynth database~\citep{RubiEtal17}. The topics, as well as
associated brain maps identified using Neurosynth, were identified and reported
in several prior studies~\citep{FoxEtal14, SulEtal17, ChenEtal20}. The topic
labels for each topic were generated automatically with the following
ChatGPT~\citep{ChatGPT} prompt: ``Please help me come up with intuitive labels
for topics I found by fitting a topic model to thousands of neuroscience and
psychology articles. I'll paste in the top 10 highest-weighted words for each
topic, and I'd like you to respond with a suggested label. For each topic,
please respond with just the topic label and no other formatting or text. Here
are the next topic's top words:'' followed by a comma-separated list of the
given topic's top-weighted words reflected in the Table~\topics. For some
topics, ChatGPT responded with a longer-form response rather than a concise
topic label. In these instances, on a case-by-case basis, we used a second
follow-up prompt to achieve the given topic's label: ``Could you please come up
with a more concise label for that topic?''. We then manually identified a set
of 11 cognitive labels that were intended to encapsulate a representative range
of widely studied low-level and high-level cognitive functions. In choosing the
set of cognitive labels, we jointly considered each topic's ChatGPT-derived
topic label, along with the top-weighted words for the topic. We attempted to
generate a concise set of labels that still spanned the full set of cognitive
functions reflected across the 80 topics. Topics that appeared unrelated to
specific cognitive functions (e.g., topics related to specific methods or
clinical themes) are designated with dashes in Table~\topics.

Finally, following an approach used in several prior studies~\citep{FoxEtal14,
SulEtal17, ChenEtal20} we treated the correlation between a given component's
brain map and each topic's brain map as an approximate measure of how much the
component was reflective of the given topic. This resulted in a set of 80
``weights'' (correlation coefficients) for each component's brain map, with one
weight per Neurosynth-derived topic.

@@

Conceptualization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
Implementation: J.R.M. and L.L.W.O. Analysis: J.R.M. and L.L.W.O. Writing,
Reviewing, and Editing: J.R.M. and L.L.W.O. Funding acquisition: J.R.M.
Supervision: J.R.M.
